---
title: "Statistical Models – Force Plate Basketball Shooting Analysis"
author: "Samuel Montalvo"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
)
knitr::opts_knit$set(root.dir = here::here())
```
# Setup and Data Loading

## Load Packages

```{r load-packages}
# Core
library(tidyverse)
library(janitor)
library(here)

# Mixed models
library(lme4)
library(lmerTest)
library(broom.mixed)

# GEE and robust SE
library(geepack)
library(sandwich)
# library(clubSandwich)  # Optional: install if needed for CR standard errors

# Model diagnostics and emmeans
library(performance)
library(emmeans)

# Tables and visualization
library(DT)
library(knitr)

# Set contrasts for proper interpretation
options(contrasts = c("contr.treatment", "contr.poly"))
```

## Helper Functions

```{r helper-functions}
# Function to find columns by keyword patterns
find_col <- function(cols, patterns, required = FALSE, name = "column") {
  for (p in patterns) {
    matches <- cols[str_detect(tolower(cols), tolower(p))]
    if (length(matches) > 0) return(matches[1])
  }
  if (required) {
    stop(paste("ERROR: Could not find required", name, "column"))
  }
  return(NA_character_)
}

# Helpers for lme4 convergence diagnostics
get_lme4_messages <- function(mod) {
  msgs <- mod@optinfo$conv$lme4$messages
  if (is.null(msgs)) character(0) else as.character(msgs)
}

is_lme4_converged <- function(messages) {
  if (length(messages) == 0) return(TRUE)
  all(str_detect(messages, "boundary \\(singular\\) fit"))
}

collapse_messages <- function(messages) {
  if (length(messages) == 0) NA_character_ else paste(unique(messages), collapse = " | ")
}

# Function to map all columns
map_columns <- function(df) {
  cols <- names(df)

  mapping <- tibble(
    role = c(
      "player_id", "shot_type", "made", "shot_number",
      "jump_height", "peak_propulsive_force", "peak_propulsive_power",
      "peak_braking_force", "peak_braking_power",
      "braking_duration", "propulsive_duration", "displacement_depth",
      "position", "height", "body_mass",
      "left_leg_propulsive", "right_leg_propulsive"
    ),
    patterns = list(
      c("^player_id$", "subject", "athlete", "^pid$"),
      c("^shot_type$", "shot.*type"),
      c("^made$", "miss.*made", "result"),
      c("shot_number", "shot_num", "trial"),
      c("jump_height", "jump"),
      c("peak_propulsive_force", "propulsive_force"),
      c("peak_propulsive_power", "propulsive_power"),
      c("peak_braking_force", "braking_force"),
      c("peak_braking_power", "braking_power"),
      c("braking_duration"),
      c("propulsive_duration"),
      c("displacement_depth", "displacement", "depth"),
      c("^position$", "position_group", "group"),
      c("^height$", "height_cm"),
      c("body_mass", "mass", "weight", "bodymass"),
      c("left.*propulsive"),
      c("right.*propulsive")
    )
  )

  mapping <- mapping %>%
    rowwise() %>%
    mutate(
      detected = find_col(cols, patterns),
      found = !is.na(detected)
    ) %>%
    ungroup()

  return(mapping)
}

# Function to create QC keep indicator
make_qc_keep <- function(df, col_map) {
  # Get column names
  brake_dur <- col_map %>% filter(role == "braking_duration") %>% pull(detected)
  prop_dur <- col_map %>% filter(role == "propulsive_duration") %>% pull(detected)
  disp_depth <- col_map %>% filter(role == "displacement_depth") %>% pull(detected)
  player_col <- col_map %>% filter(role == "player_id") %>% pull(detected)

  df <- df %>% mutate(qc_keep = TRUE, qc_reason = NA_character_)

  # Duration filters
 if (!is.na(brake_dur)) {
    df <- df %>%
      mutate(
        qc_keep = ifelse(.data[[brake_dur]] > 0.8, FALSE, qc_keep),
        qc_reason = ifelse(.data[[brake_dur]] > 0.8 & is.na(qc_reason),
                          "braking_duration > 0.8", qc_reason)
      )
  }

  if (!is.na(prop_dur)) {
    df <- df %>%
      mutate(
        qc_keep = ifelse(.data[[prop_dur]] > 0.8, FALSE, qc_keep),
        qc_reason = ifelse(.data[[prop_dur]] > 0.8 & is.na(qc_reason),
                          "propulsive_duration > 0.8", qc_reason)
      )
  }

  # Displacement depth outliers
  if (!is.na(disp_depth)) {
    # Global IQR outliers
    q1 <- quantile(df[[disp_depth]], 0.25, na.rm = TRUE)
    q3 <- quantile(df[[disp_depth]], 0.75, na.rm = TRUE)
    iqr <- q3 - q1
    lower <- q1 - 1.5 * iqr
    upper <- q3 + 1.5 * iqr

    df <- df %>%
      mutate(
        depth_iqr_outlier = .data[[disp_depth]] < lower | .data[[disp_depth]] > upper
      )

    # Within-player z-score outliers
    df <- df %>%
      group_by(.data[[player_col]]) %>%
      mutate(
        n_trials = n(),
        depth_mean = mean(.data[[disp_depth]], na.rm = TRUE),
        depth_sd = sd(.data[[disp_depth]], na.rm = TRUE),
        depth_z = ifelse(depth_sd > 0 & n_trials >= 5,
                        (.data[[disp_depth]] - depth_mean) / depth_sd, 0),
        depth_z_outlier = abs(depth_z) > 3 & n_trials >= 5
      ) %>%
      ungroup() %>%
      mutate(
        qc_keep = ifelse(depth_iqr_outlier | depth_z_outlier, FALSE, qc_keep),
        qc_reason = ifelse((depth_iqr_outlier | depth_z_outlier) & is.na(qc_reason),
                          "displacement_depth outlier", qc_reason)
      ) %>%
      select(-n_trials, -depth_mean, -depth_sd, -depth_z, -depth_iqr_outlier, -depth_z_outlier)
  }

  return(df)
}

# Function to fit LMM set for one outcome
fit_lmm_set <- function(df, outcome_col, shot_type_col, player_col,
                        covariates = NULL, data_label = "full") {

  results <- list()

  # Standardize outcome for effect size
  df <- df %>%
    mutate(
      outcome = .data[[outcome_col]],
      outcome_z = scale(outcome)[,1]
    )

  # Base model
  formula_base <- as.formula(paste("outcome ~", shot_type_col, "+ (1|", player_col, ")"))

  results$base <- tryCatch({
    lmer(formula_base, data = df, REML = TRUE)
  }, error = function(e) {
    message("Base model failed for ", outcome_col, ": ", e$message)
    NULL
  })

  # Adjusted model (if covariates available)
  if (!is.null(covariates) && length(covariates) > 0) {
    cov_string <- paste(covariates, collapse = " + ")
    formula_adj <- as.formula(paste("outcome ~", shot_type_col, "+", cov_string,
                                    "+ (1|", player_col, ")"))
    results$adjusted <- tryCatch({
      lmer(formula_adj, data = df, REML = TRUE)
    }, error = function(e) {
      message("Adjusted model failed for ", outcome_col, ": ", e$message)
      NULL
    })
  }

  # Random slope model
  formula_rs <- as.formula(paste("outcome ~", shot_type_col,
                                  "+ (1 +", shot_type_col, "|", player_col, ")"))
  results$random_slope <- tryCatch({
    mod <- lmer(formula_rs, data = df, REML = TRUE,
                control = lmerControl(optimizer = "bobyqa",
                                     optCtrl = list(maxfun = 2e5)))
    # Check for singularity
    if (isSingular(mod)) {
      message("Random slope model singular for ", outcome_col)
      attr(mod, "singular") <- TRUE
    }
    mod
  }, error = function(e) {
    message("Random slope model failed for ", outcome_col, ": ", e$message)
    NULL
  })

  # Standardized model for effect size
  formula_std <- as.formula(paste("outcome_z ~", shot_type_col, "+ (1|", player_col, ")"))
  results$standardized <- tryCatch({
    lmer(formula_std, data = df, REML = TRUE)
  }, error = function(e) NULL)

  attr(results, "outcome") <- outcome_col
  attr(results, "data_label") <- data_label
  attr(results, "n") <- nrow(df)
  attr(results, "sd_outcome") <- sd(df$outcome, na.rm = TRUE)

  return(results)
}

# Function to extract LMM results
extract_lmm_results <- function(model_list, outcome_col, data_label) {

  if (is.null(model_list$base)) {
    return(tibble(
      outcome = outcome_col,
      data = data_label,
      model = "base",
      term = NA,
      estimate = NA,
      std.error = NA,
      ci_lower = NA,
      ci_upper = NA,
      p_value = NA,
      std_beta = NA,
      icc = NA,
      var_player = NA,
      var_residual = NA,
      n = attr(model_list, "n"),
      singular = NA,
      converged = FALSE,
      convergence_note = "model fit failed"
    ))
  }

  results <- tibble()
  sd_outcome <- attr(model_list, "sd_outcome")

  for (mod_name in c("base", "adjusted", "random_slope")) {
    mod <- model_list[[mod_name]]
    if (is.null(mod)) next

    # Fixed effects
    fe <- tidy(mod, effects = "fixed", conf.int = TRUE, conf.level = 0.95)

    # Variance components
    vc <- as.data.frame(VarCorr(mod))
    var_player <- vc %>% filter(grp != "Residual") %>% pull(vcov) %>% sum()
    var_residual <- vc %>% filter(grp == "Residual") %>% pull(vcov)
    icc <- var_player / (var_player + var_residual)

    # Standardized beta from standardized model (for shot_type effect)
    std_beta <- NA
    if (!is.null(model_list$standardized) && mod_name == "base") {
      std_fe <- tidy(model_list$standardized, effects = "fixed")
      std_row <- std_fe %>% filter(str_detect(term, "shot_type|3PT"))
      if (nrow(std_row) > 0) {
        std_beta <- std_row$estimate[1]
      }
    }

    # Check singularity
    is_singular <- isSingular(mod) || isTRUE(attr(mod, "singular"))
    conv_messages <- get_lme4_messages(mod)
    did_converge <- is_lme4_converged(conv_messages)

    mod_results <- fe %>%
      mutate(
        outcome = outcome_col,
        data = data_label,
        model = mod_name,
        ci_lower = conf.low,
        ci_upper = conf.high,
        p_value = p.value,
        std_beta = ifelse(str_detect(term, "shot_type|3PT"), std_beta, NA),
        icc = icc,
        var_player = var_player,
        var_residual = var_residual,
        n = attr(model_list, "n"),
        singular = is_singular,
        converged = did_converge,
        convergence_note = collapse_messages(conv_messages)
      ) %>%
      select(outcome, data, model, term, estimate, std.error, ci_lower, ci_upper,
             p_value, std_beta, icc, var_player, var_residual, n, singular, converged,
             convergence_note)

    results <- bind_rows(results, mod_results)
  }

  return(results)
}

# Function to fit success models
fit_success_models <- function(df, made_col, shot_type_col, player_col,
                               jump_col, power_col, position_col = NULL,
                               data_label = "full") {

  results <- list()

  # Standardize predictors
  df <- df %>%
    mutate(
      made = .data[[made_col]],
      shot_type_f = factor(.data[[shot_type_col]]),
      jump_z = scale(.data[[jump_col]])[,1],
      power_z = scale(.data[[power_col]])[,1],
      player = factor(.data[[player_col]])
    )

  # Build formula
  predictors <- "shot_type_f + jump_z + power_z"
  if (!is.na(position_col) && position_col %in% names(df)) {
    df <- df %>% mutate(position_f = factor(.data[[position_col]]))
    predictors <- paste(predictors, "+ position_f")
  }

  # GEE model
  gee_formula <- as.formula(paste("made ~", predictors))

  results$gee <- tryCatch({
    geeglm(gee_formula, data = df, id = player, family = binomial,
           corstr = "exchangeable")
  }, error = function(e) {
    message("GEE model failed: ", e$message)
    NULL
  })

  # GLMM model
  glmm_formula <- as.formula(paste("made ~", predictors, "+ (1|player)"))

  results$glmm <- tryCatch({
    glmer(glmm_formula, data = df, family = binomial,
          control = glmerControl(optimizer = "bobyqa",
                                optCtrl = list(maxfun = 2e5)))
  }, error = function(e) {
    message("GLMM model failed: ", e$message)
    NULL
  })

  attr(results, "data_label") <- data_label
  attr(results, "n") <- nrow(df)
  attr(results, "n_made") <- sum(df$made, na.rm = TRUE)

  return(results)
}

# Function to extract success model results
extract_success_results <- function(model_list, data_label) {

  results <- tibble()

  # GEE results
  if (!is.null(model_list$gee)) {
    gee_summary <- summary(model_list$gee)
    gee_coef <- gee_summary$coefficients
    gee_error_code <- model_list$gee$geese$error
    gee_converged <- is.null(gee_error_code) || all(gee_error_code == 0)
    gee_note <- if (gee_converged) {
      NA_character_
    } else {
      paste("geese$error =", paste(gee_error_code, collapse = ","))
    }

    gee_results <- tibble(
      model = "GEE",
      data = data_label,
      term = rownames(gee_coef),
      estimate = gee_coef[, "Estimate"],
      std_error = gee_coef[, "Std.err"],
      or = exp(gee_coef[, "Estimate"]),
      or_lower = exp(gee_coef[, "Estimate"] - 1.96 * gee_coef[, "Std.err"]),
      or_upper = exp(gee_coef[, "Estimate"] + 1.96 * gee_coef[, "Std.err"]),
      p_value = gee_coef[, "Pr(>|W|)"],
      n = attr(model_list, "n"),
      converged = gee_converged,
      singular = NA,
      convergence_note = gee_note
    )
    results <- bind_rows(results, gee_results)
  }

  # GLMM results
  if (!is.null(model_list$glmm)) {
    glmm_fe <- tidy(model_list$glmm, effects = "fixed", conf.int = TRUE)
    glmm_messages <- get_lme4_messages(model_list$glmm)
    glmm_converged <- is_lme4_converged(glmm_messages)
    glmm_singular <- isSingular(model_list$glmm)

    glmm_results <- glmm_fe %>%
      mutate(
        model = "GLMM",
        data = data_label,
        or = exp(estimate),
        or_lower = exp(conf.low),
        or_upper = exp(conf.high),
        n = attr(model_list, "n"),
        converged = glmm_converged,
        singular = glmm_singular,
        convergence_note = collapse_messages(glmm_messages)
      ) %>%
      select(model, data, term, estimate, std.error, or, or_lower, or_upper, p.value, n,
             converged, singular, convergence_note) %>%
      rename(p_value = p.value, std_error = std.error)

    results <- bind_rows(results, glmm_results)
  }

  return(results)
}
```

## Load Data

```{r load-data}
# File paths
primary_file <- here("data", "grf_shooting_clean.csv")
fallback_file <- here("data", "GRF Shooting SM[40].csv")

# Load data
if (file.exists(primary_file)) {
  df <- read_csv(primary_file, show_col_types = FALSE)
  cat("Loaded:", primary_file, "\n")
} else if (file.exists(fallback_file)) {
  df <- read_csv(fallback_file, show_col_types = FALSE) %>%
    janitor::clean_names()
  cat("Loaded fallback:", fallback_file, "\n")
} else {
  stop("No data file found!")
}

cat("Dimensions:", nrow(df), "rows x", ncol(df), "columns\n")
```

## Column Mapping

```{r column-mapping}
col_map <- map_columns(df)

# Display mapping
col_map %>%
  filter(found) %>%
  select(role, detected) %>%
  kable(caption = "Column Mapping: Role -> Detected Column Name")
```

## Standardize Key Columns

```{r standardize-columns}
# Get detected column names
player_col <- col_map %>% filter(role == "player_id") %>% pull(detected)
shot_type_col <- col_map %>% filter(role == "shot_type") %>% pull(detected)
made_col <- col_map %>% filter(role == "made") %>% pull(detected)

# Check required columns
if (is.na(player_col)) stop("player_id column not found!")
if (is.na(shot_type_col)) stop("shot_type column not found!")

has_made <- !is.na(made_col)

# Standardize player_id
if (player_col != "player_id") {
  df <- df %>% rename(player_id = !!sym(player_col))
  player_col <- "player_id"
}
df <- df %>% mutate(player_id = as.character(player_id))

# Standardize shot_type
if (shot_type_col != "shot_type") {
  df <- df %>% rename(shot_type = !!sym(shot_type_col))
  shot_type_col <- "shot_type"
}

df <- df %>%
  mutate(
    shot_type = case_when(
      shot_type %in% c(2, "2", "2PT", "2pt") ~ "2PT",
      shot_type %in% c(3, "3", "3PT", "3pt") ~ "3PT",
      TRUE ~ as.character(shot_type)
    ),
    shot_type = factor(shot_type, levels = c("2PT", "3PT"))
  )

# Standardize made
if (has_made) {
  if (made_col != "made") {
    df <- df %>% rename(made = !!sym(made_col))
    made_col <- "made"
  }
  df <- df %>% mutate(made = as.integer(made))
}

# Update column map
col_map <- col_map %>%
  mutate(detected = case_when(
    role == "player_id" ~ "player_id",
    role == "shot_type" ~ "shot_type",
    role == "made" & has_made ~ "made",
    TRUE ~ detected
  ))

cat("\nKey columns standardized:\n")
cat("- player_id: character\n")
cat("- shot_type: factor with levels", levels(df$shot_type), "\n")
if (has_made) cat("- made: integer (0/1)\n")
```

## Data Summary

```{r data-summary}
cat("Dataset dimensions:", nrow(df), "rows x", ncol(df), "columns\n\n")

# Trials per player
trials_player <- df %>%
  count(player_id, name = "n_trials")

cat("Trials per player:\n")
print(trials_player)

# By shot type
cat("\nTrials by shot type:\n")
trials_shot <- df %>%
  count(player_id, shot_type) %>%
  pivot_wider(names_from = shot_type, values_from = n, values_fill = 0)
print(trials_shot)

# Overall counts
cat("\nOverall shot type distribution:\n")
print(table(df$shot_type))
```

# Quality Control

## Define QC Exclusions

```{r qc-definitions}
df <- make_qc_keep(df, col_map)

# Summary of exclusions
qc_summary <- df %>%
  summarise(
    total = n(),
    excluded = sum(!qc_keep),
    kept = sum(qc_keep),
    pct_excluded = round(100 * excluded / total, 1)
  )

cat("QC Summary:\n")
cat("- Total trials:", qc_summary$total, "\n")
cat("- Excluded:", qc_summary$excluded, "(", qc_summary$pct_excluded, "%)\n")
cat("- Kept:", qc_summary$kept, "\n")
```

```{r qc-by-player}
qc_by_player <- df %>%
  group_by(player_id) %>%
  summarise(
    n_total = n(),
    n_excluded = sum(!qc_keep),
    n_kept = sum(qc_keep),
    pct_excluded = round(100 * n_excluded / n_total, 1),
    .groups = "drop"
  ) %>%
  arrange(desc(pct_excluded))

cat("\nExclusions by player:\n")
datatable(qc_by_player, options = list(pageLength = 10))
```

```{r qc-by-shot-type}
qc_by_shot <- df %>%
  group_by(shot_type) %>%
  summarise(
    n_total = n(),
    n_excluded = sum(!qc_keep),
    pct_excluded = round(100 * n_excluded / n_total, 1),
    .groups = "drop"
  )

cat("\nExclusions by shot type:\n")
print(qc_by_shot)
```

```{r qc-reasons}
if (any(!df$qc_keep)) {
  cat("\nExclusion reasons:\n")
  print(table(df$qc_reason, useNA = "ifany"))
}
```

## Create Analysis Datasets

```{r create-datasets}
df_full <- df
df_qc <- df %>% filter(qc_keep)

cat("Full dataset:", nrow(df_full), "observations\n")
cat("QC dataset:", nrow(df_qc), "observations\n")
```

# Continuous Outcomes: Linear Mixed Models

## Identify Outcome Columns

```{r identify-outcomes}
# Map outcome columns
outcome_roles <- c("jump_height", "peak_propulsive_force", "peak_propulsive_power",
                   "peak_braking_force", "peak_braking_power",
                   "braking_duration", "propulsive_duration", "displacement_depth")

outcomes <- col_map %>%
  filter(role %in% outcome_roles, found) %>%
  select(role, detected)

cat("Outcome variables identified:\n")
print(outcomes)

# Primary outcomes
primary_outcomes <- outcomes %>%
  filter(role %in% c("jump_height", "peak_propulsive_force", "peak_propulsive_power",
                     "peak_braking_force", "peak_braking_power")) %>%
  pull(detected)

# Secondary outcomes (interpret cautiously)
secondary_outcomes <- outcomes %>%
  filter(role %in% c("braking_duration", "propulsive_duration", "displacement_depth")) %>%
  pull(detected)

cat("\nPrimary outcomes:", paste(primary_outcomes, collapse = ", "), "\n")
cat("Secondary outcomes:", paste(secondary_outcomes, collapse = ", "), "\n")
```

## Identify Covariates

```{r identify-covariates}
position_col <- col_map %>% filter(role == "position") %>% pull(detected)
height_col <- col_map %>% filter(role == "height") %>% pull(detected)
mass_col <- col_map %>% filter(role == "body_mass") %>% pull(detected)

covariates <- c(position_col, height_col, mass_col)
covariates <- covariates[!is.na(covariates)]

if (length(covariates) > 0) {
  cat("Covariates available:", paste(covariates, collapse = ", "), "\n")
} else {
  cat("No covariates available for adjusted models\n")
  covariates <- NULL
}
```

## Fit LMMs for All Outcomes

```{r fit-lmms, results='hide'}
# Store all results
all_lmm_results <- tibble()
all_models <- list()

# Function to fit for one dataset
fit_all_outcomes <- function(data, data_label, outcomes_vec) {
  results <- tibble()
  models <- list()

  for (outcome in outcomes_vec) {
    cat("Fitting models for", outcome, "on", data_label, "data...\n")

    model_set <- fit_lmm_set(
      df = data,
      outcome_col = outcome,
      shot_type_col = "shot_type",
      player_col = "player_id",
      covariates = covariates,
      data_label = data_label
    )

    models[[outcome]] <- model_set

    extracted <- extract_lmm_results(model_set, outcome, data_label)
    results <- bind_rows(results, extracted)
  }

  list(results = results, models = models)
}

# Fit on full data
cat("\n=== Fitting models on FULL dataset ===\n")
full_fit <- fit_all_outcomes(df_full, "full", c(primary_outcomes, secondary_outcomes))
all_lmm_results <- bind_rows(all_lmm_results, full_fit$results)
all_models$full <- full_fit$models

# Fit on QC data
cat("\n=== Fitting models on QC dataset ===\n")
qc_fit <- fit_all_outcomes(df_qc, "qc", c(primary_outcomes, secondary_outcomes))
all_lmm_results <- bind_rows(all_lmm_results, qc_fit$results)
all_models$qc <- qc_fit$models
```

## LMM Results: Shot Type Effects

```{r lmm-results-table}
# Filter to shot_type effects only
shot_type_effects <- all_lmm_results %>%
  filter(str_detect(term, "shot_type|3PT")) %>%
  filter(model == "base") %>%
  select(outcome, data, estimate, std.error, ci_lower, ci_upper, p_value, std_beta, icc, n)

cat("Shot Type (3PT vs 2PT) Effects - Base Models:\n\n")
datatable(
  shot_type_effects %>% mutate(across(where(is.numeric), ~ round(.x, 3))),
  caption = "LMM Results: 3PT vs 2PT Effect (Reference = 2PT)",
  options = list(pageLength = 20, scrollX = TRUE)
) %>%
  formatRound(columns = c("estimate", "std.error", "ci_lower", "ci_upper", "std_beta", "icc"), digits = 3) %>%
  formatSignif(columns = "p_value", digits = 3)
```

## Comparison: Full vs QC Dataset

```{r compare-full-qc}
comparison <- shot_type_effects %>%
  select(outcome, data, estimate, ci_lower, ci_upper, p_value) %>%
  pivot_wider(
    names_from = data,
    values_from = c(estimate, ci_lower, ci_upper, p_value),
    names_glue = "{data}_{.value}"
  ) %>%
  mutate(
    diff_estimate = full_estimate - qc_estimate,
    same_direction = sign(full_estimate) == sign(qc_estimate),
    both_significant = full_p_value < 0.05 & qc_p_value < 0.05
  )

cat("Comparison of Full vs QC Dataset Results:\n\n")
datatable(
  comparison %>% mutate(across(where(is.numeric), ~ round(.x, 3))),
  caption = "Full vs QC Comparison",
  options = list(scrollX = TRUE)
)
```

## Full Results Table (All Models)

```{r full-results-table}
datatable(
  all_lmm_results %>%
    filter(!is.na(term)) %>%
    mutate(across(where(is.numeric), ~ round(.x, 4))),
  caption = "Complete LMM Results",
  options = list(pageLength = 20, scrollX = TRUE),
  filter = "top"
)
```

## Model Singularity Check

```{r singularity-check}
singular_models <- all_lmm_results %>%
  filter(singular == TRUE) %>%
  distinct(outcome, data, model)

if (nrow(singular_models) > 0) {
  cat("Models with singular fit (interpret random effects cautiously):\n")
  print(singular_models)
} else {
  cat("No singular fits detected.\n")
}
```

# Binary Outcome: Shot Success Models

```{r shot-success-section, eval=has_made}
cat("Shot success models will be fitted using the 'made' variable.\n")
```

## Descriptive: Make Rates {.tabset}

```{r make-rates, eval=has_made}
# Overall
overall_rate <- mean(df_full$made, na.rm = TRUE)
cat("Overall make rate:", round(overall_rate * 100, 1), "%\n\n")

# By shot type
make_by_type <- df_full %>%
  group_by(shot_type) %>%
  summarise(
    n = n(),
    makes = sum(made),
    make_rate = round(mean(made) * 100, 1),
    .groups = "drop"
  )

cat("Make rate by shot type:\n")
print(make_by_type)

# By player
make_by_player <- df_full %>%
  group_by(player_id) %>%
  summarise(
    n = n(),
    makes = sum(made),
    make_rate = round(mean(made) * 100, 1),
    .groups = "drop"
  ) %>%
  arrange(desc(make_rate))

cat("\nMake rate by player:\n")
datatable(make_by_player, options = list(pageLength = 10))
```

## Fit Success Models

```{r fit-success-models, eval=has_made, results='hide'}
# Get column names for predictors
jump_col <- col_map %>% filter(role == "jump_height") %>% pull(detected)
power_col <- col_map %>% filter(role == "peak_propulsive_power") %>% pull(detected)
position_col <- col_map %>% filter(role == "position") %>% pull(detected)

all_success_results <- tibble()
success_models <- list()

# Full data
cat("Fitting success models on FULL data...\n")
success_full <- fit_success_models(
  df = df_full,
  made_col = "made",
  shot_type_col = "shot_type",
  player_col = "player_id",
  jump_col = jump_col,
  power_col = power_col,
  position_col = position_col,
  data_label = "full"
)
success_models$full <- success_full
all_success_results <- bind_rows(all_success_results,
                                  extract_success_results(success_full, "full"))

# QC data
cat("Fitting success models on QC data...\n")
success_qc <- fit_success_models(
  df = df_qc,
  made_col = "made",
  shot_type_col = "shot_type",
  player_col = "player_id",
  jump_col = jump_col,
  power_col = power_col,
  position_col = position_col,
  data_label = "qc"
)
success_models$qc <- success_qc
all_success_results <- bind_rows(all_success_results,
                                  extract_success_results(success_qc, "qc"))
```

## Success Model Results

```{r success-results-table, eval=has_made}
datatable(
  all_success_results %>%
    mutate(across(where(is.numeric), ~ round(.x, 3))),
  caption = "Shot Success Model Results (Odds Ratios)",
  options = list(pageLength = 20, scrollX = TRUE)
) %>%
  formatRound(columns = c("estimate", "std_error", "or", "or_lower", "or_upper"), digits = 3) %>%
  formatSignif(columns = "p_value", digits = 3)
```

## Success Model Diagnostics

```{r success-model-diagnostics, eval=has_made}
success_diagnostics <- all_success_results %>%
  distinct(model, data, n, converged, singular, convergence_note)

cat("Convergence/Singularity diagnostics for success models:\n")
print(success_diagnostics)
```

## Shot Type OR Comparison

```{r shot-type-or, eval=has_made}
shot_type_or <- all_success_results %>%
  filter(str_detect(term, "shot_type|3PT")) %>%
  select(model, data, or, or_lower, or_upper, p_value) %>%
  mutate(
    or_ci = paste0(round(or, 2), " [", round(or_lower, 2), ", ", round(or_upper, 2), "]")
  )

cat("Odds Ratios for 3PT vs 2PT:\n")
print(shot_type_or)
```

# Model Diagnostics

## Residual Diagnostics: Jump Height

```{r diagnostics-jump, fig.height=8}
jump_col <- col_map %>% filter(role == "jump_height") %>% pull(detected)

if (!is.na(jump_col) && !is.null(all_models$full[[jump_col]]$base)) {
  mod <- all_models$full[[jump_col]]$base

  # Get residuals and fitted values
  diag_data <- tibble(
    fitted = fitted(mod),
    residuals = residuals(mod),
    std_residuals = residuals(mod, scaled = TRUE)
  )

  # Residual vs Fitted
  p1 <- ggplot(diag_data, aes(x = fitted, y = residuals)) +
    geom_point(alpha = 0.5) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    geom_smooth(method = "loess", se = FALSE, color = "blue") +
    labs(title = paste("Residuals vs Fitted:", jump_col),
         x = "Fitted Values", y = "Residuals") +
    theme_minimal()

  # QQ plot
  p2 <- ggplot(diag_data, aes(sample = std_residuals)) +
    stat_qq() +
    stat_qq_line(color = "red") +
    labs(title = paste("Q-Q Plot:", jump_col),
         x = "Theoretical Quantiles", y = "Sample Quantiles") +
    theme_minimal()

  # Combine
  library(patchwork)
  combined <- p1 / p2
  print(combined)
}
```

## Residual Diagnostics: Peak Propulsive Power

```{r diagnostics-power, fig.height=8}
power_col <- col_map %>% filter(role == "peak_propulsive_power") %>% pull(detected)

if (!is.na(power_col) && !is.null(all_models$full[[power_col]]$base)) {
  mod <- all_models$full[[power_col]]$base

  diag_data <- tibble(
    fitted = fitted(mod),
    residuals = residuals(mod),
    std_residuals = residuals(mod, scaled = TRUE)
  )

  p1 <- ggplot(diag_data, aes(x = fitted, y = residuals)) +
    geom_point(alpha = 0.5) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    geom_smooth(method = "loess", se = FALSE, color = "blue") +
    labs(title = paste("Residuals vs Fitted:", power_col),
         x = "Fitted Values", y = "Residuals") +
    theme_minimal()

  p2 <- ggplot(diag_data, aes(sample = std_residuals)) +
    stat_qq() +
    stat_qq_line(color = "red") +
    labs(title = paste("Q-Q Plot:", power_col),
         x = "Theoretical Quantiles", y = "Sample Quantiles") +
    theme_minimal()

  combined <- p1 / p2
  print(combined)
}
```

## Model Issues Summary

```{r model-issues}
cat("=== Model Issues Summary ===\n\n")

# Singular fits
cat("Singular Fits:\n")
if (nrow(singular_models) > 0) {
  print(singular_models)
} else {
  cat("None\n")
}

# Convergence issues
failed_models <- all_lmm_results %>%
  filter(converged == FALSE) %>%
  distinct(outcome, data, model, convergence_note)

cat("\nFailed to Converge:\n")
if (nrow(failed_models) > 0) {
  print(failed_models)
} else {
  cat("None\n")
}
```

# Visualization of Model Results

## Estimated Marginal Means: Continuous Outcomes

```{r emmeans-plots}
emmeans_plots <- list()

for (outcome in primary_outcomes) {
  if (!is.null(all_models$full[[outcome]]$base)) {
    mod <- all_models$full[[outcome]]$base

    # Get emmeans
    emm <- emmeans(mod, ~ shot_type)
    emm_df <- as.data.frame(emm)

    p <- ggplot(emm_df, aes(x = shot_type, y = emmean, fill = shot_type)) +
      geom_col(alpha = 0.7, width = 0.6) +
      geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
      scale_fill_manual(values = c("2PT" = "#1f77b4", "3PT" = "#ff7f0e")) +
      labs(
        title = paste("Estimated Marginal Means:", outcome),
        subtitle = "Error bars = 95% CI",
        x = "Shot Type",
        y = outcome
      ) +
      theme_minimal() +
      theme(legend.position = "none")

    emmeans_plots[[outcome]] <- p
  }
}

# Display plots
for (outcome in names(emmeans_plots)) {
  cat("\n###", outcome, "\n\n")
  print(emmeans_plots[[outcome]])
}
```

## Success Model: Predicted Probabilities

```{r success-predictions, eval=has_made}
if (!is.null(success_models$full$gee)) {
  # Predict at mean covariates
  newdata <- expand.grid(
    shot_type_f = factor(c("2PT", "3PT")),
    jump_z = 0,
    power_z = 0
  )

  # Add position if in model
  if (!is.na(position_col)) {
    mode_position <- names(sort(table(df_full[[position_col]]), decreasing = TRUE))[1]
    newdata$position_f <- factor(mode_position)
  }

  # Get predictions (GEE)
  newdata$pred <- predict(success_models$full$gee, newdata = newdata, type = "response")

  p_pred <- ggplot(newdata, aes(x = shot_type_f, y = pred, fill = shot_type_f)) +
    geom_col(alpha = 0.7, width = 0.6) +
    scale_fill_manual(values = c("2PT" = "#1f77b4", "3PT" = "#ff7f0e")) +
    scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
    labs(
      title = "Predicted Make Probability by Shot Type",
      subtitle = "At mean jump height and propulsive power (GEE model)",
      x = "Shot Type",
      y = "Predicted Probability"
    ) +
    theme_minimal() +
    theme(legend.position = "none")

  print(p_pred)
}
```

```{r success-by-jump, eval=has_made}
# Probability vs jump height
if (!is.null(success_models$full$gee)) {

  # Create prediction grid
  jump_seq <- seq(-2, 2, length.out = 50)

  pred_grid <- expand.grid(
    shot_type_f = factor(c("2PT", "3PT")),
    jump_z = jump_seq,
    power_z = 0
  )

  if (!is.na(position_col)) {
    mode_position <- names(sort(table(df_full[[position_col]]), decreasing = TRUE))[1]
    pred_grid$position_f <- factor(mode_position)
  }

  pred_grid$pred <- predict(success_models$full$gee, newdata = pred_grid, type = "response")

  p_curve <- ggplot(pred_grid, aes(x = jump_z, y = pred, color = shot_type_f)) +
    geom_line(linewidth = 1.2) +
    scale_color_manual(values = c("2PT" = "#1f77b4", "3PT" = "#ff7f0e")) +
    scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
    labs(
      title = "Predicted Make Probability vs Jump Height",
      subtitle = "GEE model, at mean propulsive power",
      x = "Jump Height (z-score)",
      y = "Predicted Probability",
      color = "Shot Type"
    ) +
    theme_minimal()

  print(p_curve)
}
```

# Save Results

```{r save-results}
# Check if results folder exists
results_dir <- here("results")

if (dir.exists(results_dir)) {
  # Save LMM results
  saveRDS(all_lmm_results, file.path(results_dir, "lmm_results.rds"))
  cat("Saved: results/lmm_results.rds\n")

  # Save success model results
  if (has_made) {
    saveRDS(all_success_results, file.path(results_dir, "success_model_results.rds"))
    cat("Saved: results/success_model_results.rds\n")
  }

  # Save comparison table
  saveRDS(comparison, file.path(results_dir, "full_vs_qc_comparison.rds"))
  cat("Saved: results/full_vs_qc_comparison.rds\n")

} else {
  cat("results/ folder not found - results kept in memory only\n")
}
```

# Key Findings

```{r key-findings, results='asis'}
cat("## Summary of Key Findings\n\n")

# Extract key effects
jump_effect <- shot_type_effects %>%
  filter(str_detect(outcome, "jump"), data == "full")

power_effect <- shot_type_effects %>%
  filter(str_detect(outcome, "propulsive_power"), data == "full")

force_effect <- shot_type_effects %>%
  filter(str_detect(outcome, "propulsive_force"), data == "full")

brake_effect <- shot_type_effects %>%
  filter(str_detect(outcome, "braking_force"), data == "full")

cat("### Continuous Outcomes (3PT vs 2PT)\n\n")

if (nrow(jump_effect) > 0) {
  dir <- ifelse(jump_effect$estimate[1] > 0, "higher", "lower")
  sig <- ifelse(jump_effect$p_value[1] < 0.05, "significant", "not significant")
  cat("- **Jump Height:**", round(jump_effect$estimate[1], 2), "units", dir,
      "for 3PT (p =", format(jump_effect$p_value[1], digits = 3), ",", sig, ")\n")
}

if (nrow(power_effect) > 0) {
  dir <- ifelse(power_effect$estimate[1] > 0, "higher", "lower")
  sig <- ifelse(power_effect$p_value[1] < 0.05, "significant", "not significant")
  cat("- **Peak Propulsive Power:**", round(power_effect$estimate[1], 2), "units", dir,
      "for 3PT (p =", format(power_effect$p_value[1], digits = 3), ",", sig, ")\n")
}

if (nrow(force_effect) > 0) {
  dir <- ifelse(force_effect$estimate[1] > 0, "higher", "lower")
  sig <- ifelse(force_effect$p_value[1] < 0.05, "significant", "not significant")
  cat("- **Peak Propulsive Force:**", round(force_effect$estimate[1], 2), "units", dir,
      "for 3PT (p =", format(force_effect$p_value[1], digits = 3), ",", sig, ")\n")
}

if (nrow(brake_effect) > 0) {
  dir <- ifelse(brake_effect$estimate[1] > 0, "higher", "lower")
  sig <- ifelse(brake_effect$p_value[1] < 0.05, "significant", "not significant")
  cat("- **Peak Braking Force:**", round(brake_effect$estimate[1], 2), "units", dir,
      "for 3PT (p =", format(brake_effect$p_value[1], digits = 3), ",", sig, ")\n")
}

if (has_made) {
  cat("\n### Shot Success\n\n")

  gee_or <- all_success_results %>%
    filter(model == "GEE", data == "full", str_detect(term, "shot_type|3PT"))

  if (nrow(gee_or) > 0) {
    cat("- **3PT vs 2PT OR (GEE):**", round(gee_or$or[1], 2),
        "[", round(gee_or$or_lower[1], 2), ",", round(gee_or$or_upper[1], 2), "]",
        "(p =", format(gee_or$p_value[1], digits = 3), ")\n")
  }

  jump_or <- all_success_results %>%
    filter(model == "GEE", data == "full", str_detect(term, "jump"))

  if (nrow(jump_or) > 0) {
    cat("- **Jump Height OR (per SD):**", round(jump_or$or[1], 2),
        "[", round(jump_or$or_lower[1], 2), ",", round(jump_or$or_upper[1], 2), "]\n")
  }
}

cat("\n### QC Sensitivity Analysis\n\n")

qc_concordance <- comparison %>%
  summarise(
    same_direction = sum(same_direction, na.rm = TRUE),
    both_sig = sum(both_significant, na.rm = TRUE),
    total = n()
  )

cat("- **Direction concordance:** ", qc_concordance$same_direction, "/",
    qc_concordance$total, " outcomes have same direction in full vs QC data\n", sep = "")
cat("- **Significance concordance:** ", qc_concordance$both_sig, "/",
    qc_concordance$total, " outcomes significant in both datasets\n", sep = "")

qc_excluded_pct <- round(100 * (1 - nrow(df_qc) / nrow(df_full)), 1)
cat("- **QC exclusion rate:**", qc_excluded_pct, "%\n")

if (qc_concordance$same_direction == qc_concordance$total) {
  cat("- **Conclusion:** QC filtering does NOT materially change conclusions\n")
} else {
  cat("- **Conclusion:** Some effects differ between full and QC datasets; interpret with caution\n")
}
```

---

# Session Info

```{r session-info}
sessionInfo()
```
